Concurrent Programming with CXL

### 2020-06-16 book/paper.tex
More index items

Instead of bits and bytes, a CXL machine manipulates CXL values.
A CXL machine has the following components:
\begin{itemize}
\item Code:  This is an immutable and finite list of CXL machine instructions,
generated from a CXL program.  The types of instructions will be described later.
\item Shared memory: A CXL machine has just one memory location containing
an CXL value.
\item Processes:  Any process
can spawn an unbounded number of other processes and processes may terminate.
Each process has an immutable (but not necessarily unique) \emph{name tag},
a program counter, a stack of CXL values,
and a single mutable general purpose \emph{register} that contains a CXL value.
\end{itemize}

A name tag consists of the name of the main method of the process,
along with an optional tag specified in the \texttt{spawn} statement.
The default tag is the first argument to the method.
In Figure~\ref{fig:inc}, the created processes have name tags
\texttt{incrementer/0}, \texttt{incrementer/1}, and \texttt{main/()}.

The state of a process is called a \emph{context}: it contains the value of
its name tag, program counter, stack, and register.
The state of the CXL machine
consists of the value of its memory and the multiset (or \emph{bag}) of
contexts.  It is a multiset of contexts because two different processes can
be in the same state.
The initial state of the CXL memory is the empty dictionary, \texttt{dict\{\}}.
The context bag has an initial context in it with name tag
\texttt{\_\_main\_\_/()}, pc 0, register \texttt{()}, and an empty stack.
Each machine instruction updates the state in some way.
Figure~\ref{fig:incstate} shows an example of a reachable state for 
the program in Figure~\ref{fig:inc}.

@@

Compiling the code in Figure~\ref{fig:inc} results in the CXL machine code
listed in Figure~\ref{fig:inccode}.
You can obtain this code by invoking \texttt{cxl} with the \texttt{-a} flag
like so:
\begin{code}
\begin{verbatim}
cxl -a Up.cxl
\end{verbatim}
\end{code}
The CXL machine is predominantly a \emph{stack machine}.
All instructions are atomically executed.
Most instructions pop values from the stack or push values onto the stack.
At first there is one process (with name tag \texttt{\_\_init\_\_/()}
that initializes the state.
It starts executing at instruction 0 and keeps executing until the last
execution in the program.
In this case, the first instruction is a \texttt{JUMP} instruction that sets the
program counter to 39.
At program counter 1 is the code for the \texttt{incrementer} method.
All methods start with a \texttt{Frame} instruction and end with a \texttt{Return}
instruction.

@@

Why does it work?  We will focus here on how one might go about proving
mutual exclusion for an algorithm such as Peterson's.
For that, we have to understand a little bit more about how the CXL
machine works.
In Chapter~\ref{ch:cxlmachine} we talked about the concept of \emph{state}:
at any point of time the CXL machine is in a specific state.
A state is comprised of the values of the shared variables as well as
the values of the process variables for each process, including its
program counter and the contents of its stack.
Everytime a process executes a CXL machine instruction, the
state changes (if only because the program counter of the process
changes).  We call that a \emph{step}.
\index{step}
Steps in CXL are atomic.

@@

Here \texttt{L} is a shared list.
Another process can revive the process using the \texttt{go} stateent:

@@

Here \texttt{C} is a context and \texttt{R} is a CXL value.
It causes a process with context \texttt{C} to be added to the state that has
just executed the \texttt{stop} expression.  The \texttt{stop} expression
returns the value \texttt{R}.

>>>>

Instead of bits and bytes, a CXL machine manipulates CXL values.
A CXL machine has the following components:
\begin{itemize}
\item Code:  This is an immutable and finite list of CXL machine instructions,
generated from a CXL program.  The types of instructions will be described later.
\item Shared memory: A CXL machine has just one memory location containing
an CXL value.
\item Processes:  Any process
can spawn an unbounded number of other processes and processes may terminate.
Each process has an immutable (but not necessarily unique) \emph{name tag},
\index{name tag}
a program counter, a stack of CXL values,
and a single mutable general purpose \emph{register} that contains a CXL value.
\end{itemize}

A name tag consists of the name of the main method of the process,
along with an optional tag specified in the \texttt{spawn}
\index{spawn statement}
\index{tag}
statement.
The default tag is the first argument to the method.
In Figure~\ref{fig:inc}, the created processes have name tags
\texttt{incrementer/0}, \texttt{incrementer/1}, and \texttt{main/()}.

The state of a process is called a \emph{context} (aka continuation):
\index{context}
\index{continuation}
it contains the value of
its name tag, program counter, stack, and register.
The state of the CXL machine
consists of the value of its memory and the multiset (or \emph{bag}) of
contexts.  It is a multiset of contexts because two different processes can
be in the same state.
The initial state of the CXL memory is the empty dictionary, \texttt{dict\{\}}.
The context bag has an initial context in it with name tag
\texttt{\_\_main\_\_/()}, pc 0, register \texttt{()}, and an empty stack.
Each machine instruction updates the state in some way.
Figure~\ref{fig:incstate} shows an example of a reachable state for 
the program in Figure~\ref{fig:inc}.

@@

Compiling the code in Figure~\ref{fig:inc} results in the CXL machine code
listed in Figure~\ref{fig:inccode}.
You can obtain this code by invoking \texttt{cxl} with the \texttt{-a} flag
like so:
\begin{code}
\begin{verbatim}
cxl -a Up.cxl
\end{verbatim}
\end{code}
The CXL machine is predominantly a \emph{stack machine}.
\index{stack machine}
All instructions are atomically executed.
Most instructions pop values from the stack or push values onto the stack.
At first there is one process (with name tag \texttt{\_\_init\_\_/()}
that initializes the state.
It starts executing at instruction 0 and keeps executing until the last
execution in the program.
In this case, the first instruction is a \texttt{JUMP} instruction that sets the
program counter to 39.
At program counter 1 is the code for the \texttt{incrementer} method.
All methods start with a \texttt{Frame} instruction and end with a \texttt{Return}
instruction.

@@

Why does it work?  We will focus here on how one might go about proving
mutual exclusion for an algorithm such as Peterson's.
For that, we have to understand a little bit more about how the CXL
machine works.
In Chapter~\ref{ch:cxlmachine} we talked about the concept of \emph{state}:
\index{state}
at any point of time the CXL machine is in a specific state.
A state is comprised of the values of the shared variables as well as
the values of the process variables for each process, including its
program counter and the contents of its stack.
Everytime a process executes a CXL machine instruction, the
state changes (if only because the program counter of the process
changes).  We call that a \emph{step}.
\index{step}
Steps in CXL are atomic.

@@

Here \texttt{L} is a shared list.
Another process can revive the process using the \texttt{go}
\index{go statement}
statement:

@@

Here \texttt{C} is a context and \texttt{R} is a CXL value.
It causes a process with context \texttt{C} to be added to the state that has
just executed the \texttt{stop}
\index{stop expression}
expression.  The \texttt{stop} expression returns the value \texttt{R}.
